---
title: "ANO power analyses"
format: html
editor: visual
echo: false
---



```{r setup, include=FALSE}
options(tibble.width = Inf)
# You might also want to set tibble.print_max = Inf to display all rows
options(tibble.print_max = 50)
knitr::opts_chunk$set(message=FALSE, warning=FALSE, error=FALSE)
```

```{r}
# 
# load packages
library(tidyverse)
library(pwr)
library(lme4)
library(simr)
library(kableExtra)
```

```{r results='asis'}
#| results: 'asis'
source("C:/Users/francesca.jaroszynsk/OneDrive - NINA/nina_projects/wetlands/ano_vatmark/power_analysis.R", local = TRUE)

```

## Data structure overview

Of the 252 ANO flate sites, 38 have under the total 18 point IDs.

```{r results='asis'}
geo_ano_general |> 
  arrange(resolution, analysis_type) |> 
    kableExtra::kbl() |> #caption = paste("Summary for group:", paste(keys, collapse = " - "))
    kableExtra::kable_styling(full_width = FALSE)

```



## Power analysis results
Below are illustrated the power analysis results. Each table describes the number of plots required to detect a 1, 5 or 10% change in each of the response variables, with 60 or 80% power. Analyses have been conducted at the 1:5000 and 1:20 000 resolution, and for the 1 x 1 m quadrat analyses as well as the 250 m\^2 circle. Analyses were conducted using the function `pwr.f2.test()` from the `pwr` package.




### Results for ungrouped 'kartleggingsenheter'
```{r results='asis'}

# tables
#geo_ano_results |> 
#  filter(power == 0.8, delta_level == 10, resolution == 5000, analysis_type == "1m2") |> 
#  select(kartleggingsenhet, resolution, analysis_type, response_variable_names, n_plots) |> 
#  tidylog::left_join(geo_ano_general) |> 
#  mutate(feasibility = ifelse(n_plots < n_points, "feasible", "not feasible"))



# All vegetation types: Walk through each group and print a table
geo_ano_results_groups <- geo_ano_results |> 
  group_by(resolution, analysis_type, delta_level, power) |> 
  group_split() 
  
geo_ano_results_labels <- geo_ano_results |> 
  group_by(resolution, analysis_type, delta_level, power) |> 
  group_keys() |> 
  mutate(power = power*100)

#dat <- geo_ano_results_groups[[1]] |> as_tibble()

walk2(geo_ano_results_groups, seq_len(nrow(geo_ano_results_labels)), function(dat, idx) {
  keys <- geo_ano_results_labels[idx, ]
  
  cat("####", paste("1:", keys[1], "resolution, ", keys[2], "plot type, to detect a ", keys[3], "% change, with", keys[4], "% power."), "\n\n")
  
  summary_tbl <- dat |> 
    select(kartleggingsenhet, response_variable_names, n_plots) |> 
  pivot_wider(names_from = response_variable_names, values_from = n_plots) 
    

  print(summary_tbl  |> 
    kableExtra::kbl() |> #caption = paste("Summary for group:", paste(keys, collapse = " - "))
    kableExtra::kable_styling(full_width = FALSE) |> 
      kableExtra::kable_paper("hover")
    #kableExtra::column_spec(2, background = kableExtra::spec_color(2, option = "viridis")) |> 
    #kableExtra::kable_material_dark(c("striped", "hover"))
  )
  cat("\n\n")
})

```


### Results for grouped 'kartleggingsenheter'
Kartleggingsenheter are grouped into common and rare types.


```{r results='asis'}

# Agglomerated vegetation types: Walk through each group and print a table
geo_ano_results_groups_agglo <- geo_ano_results_agglo |> 
  group_by(resolution, analysis_type, delta_level, power) |> 
  group_split() 
  
geo_ano_results_labels_agglo <- geo_ano_results_agglo |> 
  group_by(resolution, analysis_type, delta_level, power) |> 
  group_keys() |> 
  mutate(power = power*100)


walk2(geo_ano_results_groups_agglo, seq_len(nrow(geo_ano_results_labels_agglo)), function(data, idx) {
  keys <- geo_ano_results_labels[idx, ]
  
  cat("####", paste("1:", keys[1], "resolution, ", keys[2], "plot type, to detect a ", keys[3], "% change, with", keys[4], "% power."), "\n\n")
  
  summary_tbl <- data |> 
    select(kartleggingsenhet_agglo, response_variable_names, n_plots) |> 
  pivot_wider(names_from = response_variable_names, values_from = n_plots) 
    
  
  print(summary_tbl  |> 
    kableExtra::kbl(caption = paste("Summary for group:", paste(keys, collapse = " - "))) |> 
    kableExtra::kable_styling(full_width = FALSE)
  )
  cat("\n\n")
})



```



### Figures
Graphical representation of the number of plots necessary to detect a 1, 5, or 10% change in each of the response variables at a 1:5000 or 1:20 000 resolution at 60 or 80% power, for the 1m^2 quadrats or 250m^2 circles.
```{r results='asis'}
# figures


# plotting function
plot_ano_tile <- function(geo_ano_results_figures){
  group_info <- geo_ano_results_figures |> 
    select(resolution, analysis_type, power, delta_level) |> 
    slice(1)
  
  group_label <- paste(
    paste(names(group_info), as.character(group_info), sep = "="),
    collapse = ", "
  )
  
  ggplot(geo_ano_results_figures, aes(y = kartleggingsenhet,
                                      x = response_variable_names,
                                      fill = n_plots)) +
    geom_tile() +
    scale_fill_viridis_c() +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = group_label)
}

geo_ano_results_figures |> 
  group_by(resolution, analysis_type, power, delta_level) |> 
  group_split() |> 
  walk(~ {
    if(nrow(.x) > 0) print(plot_ano_tile(.x))
  })

```
