---
title: "Appendix 1: ANO power analyses"
format: html
echo: false
---

```{r setup, include=FALSE}
options(tibble.width = Inf)
options(tibble.print_max = 50)
knitr::opts_chunk$set(message=FALSE, warning=FALSE, error=FALSE)


# function to save html tables as pngs
save_kable_as_image <- function(kable_table, filename, zoom = 2, width = 1000) {
  html_file <- tempfile(fileext = ".html")
  save_kable(kable_table, file = html_file, self_contained = TRUE)
  webshot2::webshot(html_file, file = filename, zoom = zoom, vwidth = width)
}
```

```{r}
# 
# load packages
library(tidyverse)
library(pwr)
#library(lme4)
#library(simr)
library(kableExtra)
library(flextable)
```

```{r, message=FALSE, include=FALSE}

source("C:/Users/francesca.jaroszynsk/OneDrive - NINA/nina_projects/wetlands/ano_vatmark/power_analysis.R", local = TRUE)

```

## Data structure overview

Of the 252 ANO flate sites, 38 have under the total 18 point IDs.

```{r results='asis'}
geo_ano_general |> 
  arrange(region, resolution, analysis_type) |> 
    kableExtra::kbl() |> 
    kableExtra::kable_styling(full_width = FALSE) |> 
  kableExtra::kable_paper("hover")
```

```{r}

fig <- geo_ano |> 
  ggplot(aes(x = response_variable_values)) + 
  geom_histogram() + 
  facet_wrap(~response_variable_names, scales = "free_x") + 
  theme_classic()

print(fig)

```

## Power analysis results

Below are illustrated the power analysis results. Each table describes the number of plots required to detect a 1, 5 or 10% change in each of the response variables, ~~with 60 or~~ 80% power. Analyses have been conducted at the 1:5000 and 1:20 000 resolution, and for the 1 x 1 m quadrat analyses as well as the 250 m\^2 circle. Analyses were conducted using the function `pwr.f2.test()` from the `pwr` package.

Numbers are coloured according to their reliability, whereby black values are considered reliable, and grey values unreliable. The threshold for reliability was set according to the standard deviation of each response variable and the spread of the data for each response variable (see appendix figure X):


```{r}
geo_ano_results_national |> 
  group_by(response_variable_names) |> 
  summarise(mean_sd = mean(sd_dat),
            min_sd = min(sd_dat),
            max_sd = max(sd_dat)
  ) |> 
  kableExtra::kbl() |> 
    kableExtra::kable_styling(full_width = FALSE) |> 
  kableExtra::kable_paper("hover")
  

```

The number of data points (i.e. the number of `ano_punkt_id`) per vegetation type and per response variable varies from 1 to more than 2300: 
```{r}
geo_ano_vat |> 
  group_by(grouping, response_variable_names) |> 
  summarise(n = n_distinct(ano_punkt_id)) |> 
  pivot_wider(names_from = response_variable_names, values_from = n) |> 
  kbl(escape = FALSE) |> 
  kable_paper("hover") |> 
      kable_styling(full_width = FALSE)


```



Thus, values below the standard deviation threshold are considered unreliable.

### National-level recommendations

```{r results='asis'}


# split table into 1:5000 and 1:20 000 resolution tables.
geo_ano_results_groups <- geo_ano_results_national |> 
  filter(power == 0.8) |> 
  group_by(resolution) |> 
  group_split() 
  
geo_ano_results_labels <- geo_ano_results_national |> 
  group_by(resolution) |> 
  group_keys()


walk2(geo_ano_results_groups, seq_len(nrow(geo_ano_results_labels)), function(dat, idx) {
  keys <- geo_ano_results_labels[idx, ]
  

  
  summary_tbl <- dat |> 
    mutate(hovedtype = str_extract(grouping, "^[^-]+")) |> 
    select(hovedtype, grouping, delta_level, response_variable_names, n_plots, n_obs, sd_dat, sd_threshold, obs_threshold) |>
    filter(!is.na(n_plots)) |> 
    mutate(
      obs_scale = (n_obs/1029),
      n_plots = if_else(
      (sd_dat < sd_threshold) & obs_threshold < 0.15, 
      cell_spec(n_plots, color = "grey60"),
      cell_spec(n_plots, color = "black")
    )) |> 
    ungroup() |> 
    select(-sd_dat, -sd_threshold, -n_obs, -obs_scale, -obs_threshold) |> 
  pivot_wider(names_from = response_variable_names, values_from = n_plots) |> 
    arrange(hovedtype, delta_level) |> 
    select(-hovedtype)
    

  print(summary_tbl  |> 
    kbl(escape = FALSE, caption = paste("1:", keys[1], "resolution")) |> 
      kable_paper() |> 
      add_header_above(c(" ", "1m^2 plot" = 6, "250m^2 circle" = 2), line_sep = 5) |> 
      kable_styling(full_width = FALSE)

      )
})

```


```{r}
geo_ano_results_national |> 
  mutate(n_observations = paste0(n_obs, "/", total_obs)) |> 
  distinct(grouping, resolution, response_variable_names, n_observations) |> 
  pivot_wider(names_from = response_variable_names, values_from = n_observations) |> 
  kbl(escape = FALSE) |> 
  kable_paper() |> 
      kable_styling(full_width = FALSE)

```



### Geopolitical regional-level recommendations

```{r results='asis'}


# split table into 1:5000 and 1:20 000 resolution tables.
geo_ano_results_groups <- geo_ano_results_regional |> 
  filter(power == 0.8) |> 
  group_by(resolution) |> 
  group_split() 
  
geo_ano_results_labels <- geo_ano_results_regional |> 
  group_by(resolution) |> 
  group_keys() #|> 
  #mutate(power = power*100)


walk2(geo_ano_results_groups, seq_len(nrow(geo_ano_results_labels)), function(dat, idx) {
  keys <- geo_ano_results_labels[idx, ]
  

  
  summary_tbl <- dat |> 
    mutate(hovedtype = str_extract(grouping, "^[^-]+")) |> 
    select(region, hovedtype, grouping, delta_level, response_variable_names, n_plots, n_obs, sd_dat, sd_threshold) |>
    filter(!is.na(n_plots)) |> 
    mutate(
      obs_scale = (n_obs/1029),
      n_plots = if_else(
      (sd_dat < sd_threshold) & (obs_scale < 0.15), 
      cell_spec(n_plots, color = "grey60"),
      cell_spec(n_plots, color = "black")
    )) |> 
    ungroup() |> 
    select(-sd_dat, -sd_threshold, -n_obs, -obs_scale) |> 
  pivot_wider(names_from = response_variable_names, values_from = n_plots) |> 
    arrange(hovedtype, delta_level) |> 
    select(-hovedtype)
    

  print(summary_tbl  |> 
    kbl(escape = FALSE, caption = paste("1:", keys[1], "resolution")) |> #, ", keys[2], "plot type" , to detect a ", keys[3], "% change, with", keys[4], "% power."     #caption = paste("Summary for group:", paste(keys, collapse = " - "))
      kable_paper() |> 
      add_header_above(c(" " = 2, "1m^2 plot" = 6, "250m^2 circle" = 2), line_sep = 5) |> 
      kable_styling(full_width = FALSE)

      )
})

```


### Bioclimatice regional-level recommendations

```{r results='asis'}

# split table into 1:5000 and 1:20 000 resolution tables.
geo_ano_results_groups <- geo_ano_results_bioclimatic |> 
  filter(power == 0.8) |> 
  group_by(resolution) |> 
  group_split() 
  
geo_ano_results_labels <- geo_ano_results_bioclimatic |> 
  group_by(resolution) |> 
  group_keys() #|> 
  #mutate(power = power*100)

#dat <- geo_ano_results_groups[[1]] |> as_tibble()

walk2(geo_ano_results_groups, seq_len(nrow(geo_ano_results_labels)), function(dat, idx) {
  keys <- geo_ano_results_labels[idx, ]
  

  
  summary_tbl <- dat |> 
    mutate(hovedtype = str_extract(grouping, "^[^-]+")) |> 
    select(BCregion, hovedtype, grouping, delta_level, response_variable_names, n_plots, n_obs, sd_dat, sd_threshold) |>
    filter(!is.na(n_plots)) |> 
    mutate(
      obs_scale = (n_obs/1029),
      n_plots = if_else(
      (sd_dat < sd_threshold) & (obs_scale < 0.15), 
      cell_spec(n_plots, color = "grey60"),
      cell_spec(n_plots, color = "black")
    )) |> 
    ungroup() |> 
    select(-sd_dat, -sd_threshold, -n_obs, -obs_scale) |> 
  pivot_wider(names_from = response_variable_names, values_from = n_plots) |> 
    arrange(hovedtype, delta_level) |> 
    select(-hovedtype)
    

  print(summary_tbl  |> 
    kbl(escape = FALSE, caption = paste("1:", keys[1], "resolution")) |> #, ", keys[2], "plot type" , to detect a ", keys[3], "% change, with", keys[4], "% power."     #caption = paste("Summary for group:", paste(keys, collapse = " - "))
      kable_paper() |> 
      add_header_above(c(" " = 2, "1m^2 plot" = 6, "250m^2 circle" = 2), line_sep = 5) |> 
      kable_styling(full_width = FALSE)

      )
})

```


