---
title: "Appendix 1: ANO power analyses"
format: 
  html:
    embed-resources: true
    number-sections: true
echo: false
---

```{r setup, include=FALSE}
options(tibble.width = Inf)
options(tibble.print_max = 50)
knitr::opts_chunk$set(message=FALSE, warning=FALSE, error=FALSE)


# function to save html tables as pngs
save_kable_as_image <- function(kable_table, filename, zoom = 2, width = 1000) {
  html_file <- tempfile(fileext = ".html")
  save_kable(kable_table, file = html_file, self_contained = TRUE)
  webshot2::webshot(html_file, file = filename, zoom = zoom, vwidth = width)
}
```

```{r}
# 
# load packages
library(tidyverse)
library(pwr)
#library(lme4)
#library(simr)
library(kableExtra)
library(flextable)
```

```{r, message=FALSE, include=FALSE}

source("C:/Users/francesca.jaroszynsk/OneDrive - NINA/nina_projects/wetlands/ano_vatmark/power_analysis.R", local = TRUE)

```

## Methods

A threshold for reliability was set according to the standard deviation (SD) of each response variable and the spread of the data for each response variable, whereby the recommended number of plots was considered unreliable when the SD was lower than 10% of the range, and where the number of observations was lower than 15% of the total number of observations for each explanatory variable, or where there are simply too few observations (less than 5). Analyses have been conducted at the 1:5000 and 1:20 000 resolution, and for the 1 x 1 m quadrat analyses as well as the 250 m\^2 circle. Analyses were conducted using the function `pwr.f2.test()` from the `pwr` package. Analyses were performed using R Statistical Software (`r R.version.string`, RCoreTeam).

```{r results='asis'}
geo_ano_general |> 
  select(-median_n_plots_site) |> 
  arrange(region) |>
  kbl(caption = "Table: Overview of the number of plots per region and vegetation type.") |> 
  kable_styling(full_width = FALSE)
```

```{r}
#| fig-width: 8
#| fig-height: 7
#| out-width: "100%"

fig <- geo_ano |> 
  ggplot(aes(x = response_variable_values)) + 
  geom_histogram() + 
  facet_wrap(~response_variable_names, scales = "free") + 
  theme_classic()

print(fig)

```


```{r results='asis'}
geo_ano_results |> 
  group_by(response_variable_names) |> 
  summarise(mean_sd = mean(sd_dat_mean),
            min_sd = min(sd_dat_mean),
            max_sd = max(sd_dat_mean)
  ) |> 
  kableExtra::kbl(caption = "Table: ") |> 
  kableExtra::kable_styling(full_width = FALSE) |> 
  kableExtra::kable_paper("hover")
  

```

The number of data points (i.e. the number of `ano_punkt_id`) per vegetation type and per response variable varies from 1 to more than 2300:

```{r results='asis'}
geo_ano_vat |> 
  group_by(grouping, response_variable_names) |> 
  summarise(n = n_distinct(ano_punkt_id)) |> 
  pivot_wider(names_from = response_variable_names, values_from = n) |> 
  kbl() |> 
  kable_paper("hover")


```


## Results

### Detectability
The following tables highlight the detectability of wetland vegetation types (V1-V9) within the ANO database.

#### National
```{r results='asis'}

site_detectability |> 
  #filter(grouping_value == "national") |> 
  kbl() |> 
  kable_paper("hover") |> 
  kable_styling(full_width = FALSE)

```

##### Geopolitical
#```{r results='asis'}
#
#site_detectability_all |> 
#  filter(grepl("Norway", grouping_value)) |> 
#  kbl() |> 
#  kable_paper("hover") |> 
#  kable_styling(full_width = FALSE)
#
#```
#
##### Bioclimatic
#```{r results='asis'}
#
#site_detectability_all |> 
#  filter(!grepl("Norway", grouping_value), grouping_value!= "national") |> 
#  kbl() |> 
#  kable_paper("hover") |> 
#  kable_styling(full_width = FALSE)
#
#```



### Power analysis recommendations

Below are illustrated the power analysis results. Each table describes the number of plots required to detect a 1 or 5% change in each of the response variables, with ~~60 or~~ 80% power. The number of plots required to achieve this are coloured according to their reliability, whereby black values are considered reliable, and grey values unreliable.

#### National

```{r results='asis'}


# split table into 1:5000 and 1:20 000 resolution tables.
geo_ano_results_groups <- geo_ano_results |> 
  #filter(power == 0.8) |> 
  group_by(power) |> 
  group_split() 
  
geo_ano_results_labels <- geo_ano_results |> 
  group_by(power) |> 
  group_keys()


walk2(geo_ano_results_groups, seq_len(nrow(geo_ano_results_labels)), function(dat, idx) {
  keys <- geo_ano_results_labels[idx, ]
  

  
  summary_tbl <- dat |> 
    mutate(hovedtype = str_extract(grouping, "^[^-]+")) |> 
    select(hovedtype, grouping, delta_level, response_variable_names, n_plots, n_obs, sd_dat_mean, sd_threshold, obs_threshold) |>
    filter(!is.na(n_plots),
           # select only 1 and 5% delta levels
           delta_level %in% c("1", "5")) |> 
    mutate(
      n_plots = if_else(
      ((sd_dat_mean < sd_threshold) & (obs_threshold < 0.15))|(n_obs <= 3), 
      cell_spec(n_plots, color = "grey60"),
      cell_spec(n_plots, color = "black")
    )) |> 
    ungroup() |> 
    select(-sd_dat_mean, -sd_threshold, -n_obs, -obs_threshold) |> 
  pivot_wider(names_from = response_variable_names, values_from = n_plots) |> 
    arrange(hovedtype, delta_level) |> 
    select(-hovedtype)
    

  print(summary_tbl  |> 
    kbl(escape = FALSE) |> #, caption = paste("1:", keys[1], "resolution") 
      kable_paper("hover") #|> 
      #add_header_above(c(" ", "1m^2 plot" = 6, "250m^2 circle" = 1), line_sep = 5) |> 
      #column_spec(1:2, extra_css = "white-space: nowrap;")
      )
})

```



# #### Geographical
#
#```{r results='asis'}
#
#
## split table into 1:5000 and 1:20 000 resolution tables.
#geo_ano_results_groups <- geo_ano_results_regional |> 
#  filter(power == 0.8) |> 
#  group_by(resolution) |> 
#  group_split() 
#  
#geo_ano_results_labels <- geo_ano_results_regional |> 
#  group_by(resolution) |> 
#  group_keys() #|> 
#  #mutate(power = power*100)
#
#
#walk2(geo_ano_results_groups, seq_len(nrow(geo_ano_results_labels)), function(dat, idx) {
#  keys <- geo_ano_results_labels[idx, ]
#  
#
#  
#  summary_tbl <- dat |> 
#    mutate(hovedtype = str_extract(grouping, "^[^-]+")) |> 
#    select(Region, hovedtype, grouping, delta_level, response_variable_names, n_plots, n_obs, sd_dat, sd_threshold, obs_threshold) |>
#    filter(!is.na(n_plots),
#           # select only 1 and 5% delta levels
#           delta_level %in% c("1", "5")) |> 
#    mutate(
#      n_plots = if_else(
#      ((sd_dat < sd_threshold) & (obs_threshold < 0.15))|(n_obs <= 3), 
#      cell_spec(n_plots, color = "grey60"),
#      cell_spec(n_plots, color = "black")
#    )) |> 
#    ungroup() |> 
#    select(-sd_dat, -sd_threshold, -n_obs, -obs_threshold) |> 
#  pivot_wider(names_from = response_variable_names, values_from = n_plots) |> 
#    arrange(hovedtype, delta_level) |> 
#    select(-hovedtype)
#    
#
#  print(summary_tbl  |> 
#    kbl(escape = FALSE, caption = paste("1:", keys[1], "resolution")) |> #, ", keys[2], "plot type" , to detect a ", keys[3], "% change, with", keys[4], "% power."     ##caption = paste("Summary for group:", paste(keys, collapse = " - "))
#      kable_paper("hover") |> 
#      add_header_above(c(" " = 2, "1m^2 plot" = 6, "250m^2 circle" = 1), line_sep = 5) |> 
#      kable_styling(full_width = FALSE) |> 
#      column_spec(1:2, extra_css = "white-space: nowrap;") |> 
#      column_spec(8, border_right = "1px solid grey")
#      )
#})
#
#```
#
##### Bioclimatic
#
#```{r results='asis'}
#
## split table into 1:5000 and 1:20 000 resolution tables.
#geo_ano_results_groups <- geo_ano_results_bioclimatic |> 
#  filter(power == 0.8) |> 
#  group_by(resolution) |> 
#  group_split() 
#  
#geo_ano_results_labels <- geo_ano_results_bioclimatic |> 
#  group_by(resolution) |> 
#  group_keys() #|> 
#  #mutate(power = power*100)
#
##dat <- geo_ano_results_groups[[1]] |> as_tibble()
#
#walk2(geo_ano_results_groups, seq_len(nrow(geo_ano_results_labels)), function(dat, idx) {
#  keys <- geo_ano_results_labels[idx, ]
#  
#
#  
#  summary_tbl <- dat |> 
#    mutate(hovedtype = str_extract(grouping, "^[^-]+")) |> 
#    select(Bioclimatic_region, hovedtype, grouping, delta_level, response_variable_names, n_plots, n_obs, sd_dat, sd_threshold, obs_threshold) |>
#    filter(!is.na(n_plots),
#           # select only 1 and 5% delta levels
#           delta_level %in% c("1", "5")) |> 
#    mutate(
#      n_plots = if_else(
#      ((sd_dat < sd_threshold) & (obs_threshold < 0.15))|(n_obs <= 3), 
#      cell_spec(n_plots, color = "grey60"),
#      cell_spec(n_plots, color = "black")
#    )) |> 
#    ungroup() |> 
#    select(-sd_dat, -sd_threshold, -n_obs, -obs_threshold) |> 
#  pivot_wider(names_from = response_variable_names, values_from = n_plots) |> 
#    arrange(hovedtype, delta_level) |> 
#    select(-hovedtype)
#    
#
#  print(summary_tbl  |> 
#    kbl(escape = FALSE, caption = paste("1:", keys[1], "resolution")) |> #, ", keys[2], "plot type" , to detect a ", keys[3], "% change, with", keys[4], "% power."     ##caption = paste("Summary for group:", paste(keys, collapse = " - "))
#      kable_paper("hover") |> 
#      add_header_above(c(" " = 2, "1m^2 plot" = 6, "250m^2 circle" = 1), line_sep = 5) |> 
#      kable_styling(full_width = FALSE) |> 
#      column_spec(1:2, extra_css = "white-space: nowrap;") |> 
#      column_spec(8, border_right = "1px solid grey")
#
#      )
#})
#
#```
#


### Unused tables

```{r results='asis'}
geo_ano_results |> 
  mutate(n_observations = paste0(n_obs, " / ", total_obs)) |> 
  distinct(grouping, response_variable_names, n_observations) |> 
  pivot_wider(names_from = response_variable_names, values_from = n_observations) |> 
  kbl(escape = FALSE) |> 
  kable_paper("hover") |> 
      kable_styling(full_width = FALSE)

```
